---
title: "How does your income compare?"
description: |
  An interactive gadget to compare yourself to other Australians.
author:
  - name: Nora Jones
    url: https://example.com/norajones
date: 03-23-2018
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Have you ever wondered how your income compares to other Australians? Well use this calculator thing I made to look it up then. It takes your household's after-tax income and adjusts it for the number of people in your house to give your 'equivalised disposable household income.' The calculator uses <a href="http://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/6523.02015-16?OpenDocument">ABS data from 2015-16</a>, which was released in September 2017.

```{r}
knitr::include_app('https://mattcowgill.shinyapps.io/income_dist/', height = '1100px')
```


You can <a href="http://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/6523.02015-16?OpenDocument">get the ABS data here</a>.